---
title: "Adv_R_Homework1"
author: "Zixuan Gao 23206703"
format: pdf
editor: visual
execute: 
  echo: false
  warning: false
  format:
  pdf:
    toc: true
    toc-title: \textcolor{Orange}{Contents}
    embed-resources: true
    fontsize: 1.2em
---

## \textcolor{Orange}{1. Introduction}

The **'Orange'** is a dataset which is available in R. This dataset has 35 rows and 3 columns of records of the growth of orange trees. This dataset was originally part of package **nlme**, and that has methods *(including for \[, as.data.frame, plot and print)* for its grouped-data classes.It is extracted from Draper, N. R. and Smith, H. (1998), *Applied Regression Analysis (3rd ed)*, Wiley (exercise 24.N) & Pinheiro, J. C. and Bates, D. M. (2000) *Mixed-effects Models in S and S-PLUS*, Springer.(See [Orange](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/Orange.html) for more details).

This analysis explores the **relationship between orange tree age and circumference.**

## \textcolor{Orange}{Data loading}

Below the first 6 rows of the dataset are displayed:

```{r}
data("Orange")
head(Orange)
```

### \textcolor{Orange}{Description}

Below are some descriptive statistics of the tree age and circumference in **Orange** dataset.

```{r}
summary(Orange)
```

## \textcolor{Orange}{2. Data visualisation}

We can plot the data and visually analyze it using the R package *ggplot2*:

```{r}
#| label: plot
#| fig-cap: "Relationship between OrangeTree Age and Circumference in Orange dataset"
#| fig-width: 8
#| fig-height: 6
#| warning: false
#| echo: false

# Load the necessary libraries
library(ggplot2)

# Create the scatter plot
ggplot(Orange, aes(x=age, y=circumference)) +  geom_point() +  geom_line(col = 'orange') +
  labs(
    title = "Relationship between OrangeTree Age and Circumference",
    x = "Age",
    y = "Circumference"  ) +  
  theme_gray() +  theme(plot.title = element_text(hjust = 0.5))
```

## \textcolor{Orange}{3. Summary}

From the line graph, we can see that the age of the orange tree is positively correlated with its circumference. The older the tree is, the larger its circumference is. However, as the tree ages, the differences in circumference also increase. Trees of the same age may have very different circumferences. For example, when the tree is close to 1600 years old, its circumference is less than 150 at the minimum and greater than 220 at the maximum.
